The value of being indispensable: a cooperative approach to bargaining with a continuum of players

### 2023-11-21 src/paper/theory.tex
Reorganize many-sided case

Up until this point, we have considered games with a single type of small players.
Now imagine that, in addition to player $P$, there are $L$ types of smaller players: $\{A^l_1, \dots, A^l_n\}$ where $1 \leq l \leq L$.
As before, assume that $P$ is necessary for any coalition to have a positive value, and the small players are identical to each other within their types.
Formally, the game in coalitional form is now the following: the set of players is $N = \{P, A^1_1, \dots, A^1_n, \dots, A^L_1, \dots, A^L_n\}$ and the characteristic function is
\begin{align*}
    v(S) = \begin{cases}
        0                                                & \text{if } P \notin S \\
        f\left(\frac{n_{A^1}(S)}{n}, \dots, \frac{n_{A^L}(S)}{n}\right) & \text{otherwise}.
    \end{cases}
\end{align*}

A prominent interpretation of the two-sided version of this model would be a platform marketplace with a set of sellers and a set of buyers, where all three sides possess some amount of bargaining power.
However, player $P$ does not have to be in the ``middle'' of the transactions for this framework to be applicable.
For example, it also captures the situation of a single upstream producer, a large number of downstream firms, and a similarly large number of customers, as long as both the customers and the downstream firms can participate in the bargaining process.
A setting where these assumptions might be plausible is, for example, the new car market with a car producer, a number of independent dealerships, and customers who might engage in some form of negotiation.

Let us follow the pattern from the previous sections.
Start by characterizing conditions for monotonicity and superadditivity to support the bargaining interpretation and the formation of the grand coalition.
The conditions are the straightforward analogues of those from the one-sided case.
\begin{proposition}
    The game $(N, v)$ is monotone and superadditive if and only if $f$ is increasing in all of its arguments.
\end{proposition}

Next, let us obtain formulas for the values of the three types of players.
It turns out that the limits of the Shapley-values (as $n \to \infty$) are still analytically tractable, and they also have a neat interpretation.
\begin{proposition}
    Let $f$ be continuous everywhere on $[0, 1]^L$.
    Then,
    \begin{align*}
        \varphi_P^\infty & = \int_0^1 f(t, \dots, t) \dt                                 \\
        \varphi_{A^l}^\infty & = \int_0^1 t \partial_l f(t, \dots, t) \dt.
    \end{align*}
\end{proposition}

As the previous proposition shows, the value of each kind of player depends on their expected marginal contribution, i.e. the partial derivatives of $f$.
Consequently, the value of the grand coalition can be decomposed into the Shapley-values of its members the following way:
\begin{align*}
    f(1, \dots, 1) = \underbrace{\int_0^1 f(t, \dots, t) \dt}_{\text{ player } P} + \sum_{l=1}^L \underbrace{\int_0^1 t \partial_l f(t, \dots, t) \dt}_{\text{players } A^ll}.
\end{align*}
This formulation makes it clear that values are intimately related to the marginal contributions of the players, captured by the partial derivatives of $f$ in the case of the various fringe players.

One surprising element of the result is that one only has to integrate over the diagonal of the support  of $f$, and the value does not depend on off-diagonal parts of the $f$ fucntion.
The intuition for that relies on a law-of-large-numbers-type reasoning.
If the number of fringe players is large, then it is very unlikely that, after a random ordering, their proportion is too different on any interval.
Consequently, when taking the expectation over all orderings, those with very unequal numbers of the various types have little impact on it.

@@

    % Notice that, regardless of $s$, the variance $\Var(X_n | S_n = s)$ is bounded by $\frac{1}{n}$.
    % By Chebyshev's inequality,
    % \begin{align*}
    %     \Pr( \| X_n - (s, \dots, s) \| > \epsilon \mid S_n = s) \leq \frac{L}{n \epsilon^2}.
    % \end{align*}

@@

\section{Proofs of propositions and lemmas in the main text}
\label{sec:proofs}

>>>>

Up until this point, we have only considered games with a single type of small player.
Now imagine that, in addition to player $P$, there are $L$ types of smaller players: $\{A^l_1, \dots, A^l_n\}$ where $1 \leq l \leq L$.
As before, assume that $P$ is necessary for any coalition to have a positive value, and the small players are identical to each other \emph{within their types}.
Formally, the game in coalitional form is now the following: the set of players is $N = \{P, A^1_1, \dots, A^1_n, \dots, A^L_1, \dots, A^L_n\}$ and the characteristic function is
\begin{align*}
    v(S) = \begin{cases}
        0                                                & \text{if } P \notin S \\
        f\left(\frac{n_{A^1}(S)}{n}, \dots, \frac{n_{A^L}(S)}{n}\right) & \text{otherwise}.
    \end{cases}
\end{align*}

A prominent interpretation of the two-sided version of this model would be a platform marketplace with a set of sellers and a set of buyers, where all three sides possess some amount of bargaining power.
However, player $P$ does not have to be in the ``middle'' of the transactions for this framework to be applicable.
For example, it also captures the situation of a single upstream producer, a large number of downstream firms, and a similarly large number of customers, as long as both the customers and the downstream firms can participate in the bargaining process.\footnote{
    A setting where these assumptions might be plausible is, for example, the new car market with a car producer, a number of independent dealerships, and customers who might engage in some form of negotiation.
}
Or, following \textcite{stole1996intra}, it could also be used to model bargaining between a firm and multiple types of workers.

Let us follow the pattern from the previous sections.
Start by characterizing conditions for monotonicity and superadditivity to support the bargaining interpretation and the formation of the grand coalition.
It is easy to see that the conditions are the straightforward analogues of those from the one-sided case.
\begin{proposition}
    The game $(N, v)$ is monotone and superadditive if and only if $f$ is increasing in all of its arguments.
\end{proposition}

In the next subsections, I will first examine random order values, and then two important special cases: the Shapley value and the weighted value.

% \subsection{Shapley value}

% Let us obtain formulas for the values of the three types of players.
% It turns out that the limits of the Shapley-values (as $n \to \infty$) are still analytically tractable, and they also have a neat interpretation.
% \begin{proposition}
%     Let $f$ be continuous everywhere on $[0, 1]^L$.
%     Then,
%     \begin{align*}
%         \varphi_P^\infty & = \int_0^1 f(t, \dots, t) \dt                                 \\
%         \varphi_{A^l}^\infty & = \int_0^1 t \partial_l f(t, \dots, t) \dt.
%     \end{align*}
% \end{proposition}

% As the previous proposition shows, the value of each kind of player depends on their expected marginal contribution, i.e. the partial derivatives of $f$.
% Consequently, the value of the grand coalition can be decomposed into the Shapley-values of its members the following way:
% \begin{align*}
%     f(1, \dots, 1) = \underbrace{\int_0^1 f(t, \dots, t) \dt}_{\text{ player } P} + \sum_{l=1}^L \underbrace{\int_0^1 t \partial_l f(t, \dots, t) \dt}_{\text{players } A^ll}.
% \end{align*}
% This formulation makes it clear that values are intimately related to the marginal contributions of the players, captured by the partial derivatives of $f$ in the case of the various fringe players.

% One surprising element of the result is that one only has to integrate over the diagonal of the support  of $f$, and the value does not depend on off-diagonal parts of the $f$ fucntion.
% The intuition for that relies on a law-of-large-numbers-type reasoning.
% If the number of fringe players is large, then it is very unlikely that, after a random ordering, their proportion is too different on any interval.
% Consequently, when taking the expectation over all orderings, those with very unequal numbers of the various types have little impact on it.

@@

@@

\section{Proofs of propositions in the main text}
\label{sec:proofs}
